
\documentclass[11pt,a4paper]{report}		                                   

\usepackage{graphicx}
\usepackage{a4}
\usepackage[english, ngerman]{babel}
%That Headings have Kapitälchen:
\usepackage[T1]{fontenc}

\usepackage{xcolor}

%%%%%%%%%BIBLIOGRAPHY + PICTURES IN TABLE OF CONTENTS
\usepackage[nottoc]{tocbibind}
\usepackage{breakcites}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%ABKÜRZUNGEN
\usepackage{nomencl}
\let\abbrev\nomenclature
\renewcommand{\nomname}{List of Abbreviations}
\setlength{\nomlabelwidth}{.25\hsize}
\renewcommand{\nomlabel}[1]{#1 \dotfill}
\setlength{\nomitemsep}{-\parsep}
\makenomenclature 
\newcommand{\Abkuerzung}{
    \printnomenclature
    \newpage
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%Links
\usepackage{hyperref}
\usepackage{url}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%Listings
\usepackage{listings} 

\lstdefinestyle{XQuery}{
	language=XQuery,
	float=h,
}

\xdefinecolor{listing-blue}{rgb}{33,43,63}
%\xdefinecolor{listing-green}{rgb}{00A000}
%\xdefinecolor{listing-red}{rgb}{D00000}
%\xdefinecolor{listing-cyan}{rgb}{00A0A0}
%\xdefinecolor{listing-purple}{rgb}{AE26AE}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%XQUERY LANGUAGE
\lstdefinelanguage{XQuery}{
	keywordsprefix=\$,
	keywords={lalalalala},
	morekeywords=[2]{for,let,where,order by,return,ascending,descending,:=,mod,empty in,window,group,toi,fn:data },
%	morekeywords={ancestor,ancestor-or-self,and,as,ascending,at,attribute,base-uri,boundary-space,by,case,cast,castable,
%		child,collation,comment,construction,copy-namespaces,declare,descendant,descendant-or-self,descending,
%		div,document,document-node,element,else,empty,empty-sequence,encoding,eq,every,except,external,following,
%		following-sibling,for,function,ge,greatest,gt,idiv,if,import,in,inherit,instance,intersect,item,lax,le,
%		least,let,lt,mod,module,namespace,ne,node,no-inherit,no-preserve,option,or,order,ordered,ordering,
%		preceding,preceding-sibling,preserve,processing-instruction,return,satisfies,schema,schema-attribute,schema-element,
%		self,some,stable,strict,strip,text,then,treat,typeswitch,union,unordered,validate,variable,version,where,xquery},
	% removed from the list of keywords: parent,is,of,to,default
%	otherkeywords={-,;,::,:=,!=,?,?>,/,//,/>,.,..,(\#,[,],]]>,\{,\},@,\$,*,\#),+,<,<!--,<![CDATA[,<?,</,<<,<=,=,>,-->,>=,>>,|},
%	sensitive,
%	morecomment=[s]{(:}{:)},
	string=[b]",
	morestring=[s]{'}{'},
	morestring=[s]{"`}{"'},
%	morestring=[b]",
%	morestring=[b]',
}

\lstset{
   language=XQuery,
   basicstyle=\footnotesize,
%   backgroundcolor=\color{yellow},
   keywordstyle=\color{yellow},
   keywordstyle=[1]\color{green},
   keywordstyle=[2]\color{blue}, 
   keywordsprefix=\$,
   identifierstyle=,
   commentstyle=\color{green!50!black!100},
   stringstyle=\color{red!100},
   breaklines=true,
   numbers=left,
   tabsize=2,
   numberstyle=\footnotesize,
   frame=single
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%GNUPLOT
\usepackage{gnuplottex}   
%\usepackage{epstopdf}

\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%QUERTABELLEN
\usepackage{rotating} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%TABELLEN
\usepackage{multirow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%SCHUSTERJUNGEN UND HURENKINDER
\clubpenalty10000
\widowpenalty10000
\displaywidowpenalty=10000

%%%%%%%%%%%%%%%%%%%%%%%%%%%KEIN Ba sex
\selectlanguage{english}
\hyphenation{BaseX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%APPENDIX
\usepackage[toc,page]{appendix}



\newcommand{\thema}{Migrating the XML Database \textsc{BaseX} to the Android Platform and Analyzing the Result}
\newcommand{\schlagworte}{\textsc{BaseX}, XML, Database, Android, Performance, SQLite3}
\newcommand{\zusammenfassung}{
The present thesis outlines the migration of the XML database \textsc{BaseX} to the Android platform.
\textsc{BaseX} is available for desktop platforms, therefore it is outlined how Android and those platforms are differ.
All necessary steps to build \textsc{BaseX} on Android are described and the process of creating an Android library which offers all needed \textsc{BaseX} functionalities is shown.
After having created the working Android library of \textsc{BaseX}, the performance of the database has been analyzed and improved.
For this purpose, different benchmarks on various devices have been used to identify the bottlenecks of the mobile \textsc{BaseX} version.
After identifying these problems, it is shown which of them are depending on a specific platform constraint and which can be improved by changing the software.
The performance differences between \textsc{BaseX} and the default Android database SQLite3 are also illustrated and which of these should be prefered for a specific domain.

}
\newcommand{\ausgabedatum}{28.09.2013}
\newcommand{\abgabedatum}{28.03.2014}
\newcommand{\autor}{Stephan Hagios}
\newcommand{\autorStrasse}{Zasiusstra\ss e 17}
\newcommand{\autorPLZ}{78462}
\newcommand{\autorOrt}{ Konstanz}
\newcommand{\autorGeburtsort}{Freiburg, im Breisgau}
\newcommand{\autorGeburtsdatum}{03.07.1985 }
\newcommand{\prueferA}{Prof. Dr. Oliver Eck}
\newcommand{\prueferB}{Dr. Christian Gr\"un}
\newcommand{\prueferC}{Dr. Alexander Holupirek}
\newcommand{\firma}{\textsc{BaseX} GmbH}
\newcommand{\studiengang}{Master of Science Informatik}


\begin{document}

\include{cover}
\include{title}
\include{abstract}
\include{affidavit}

\selectlanguage{english}
\tableofcontents
\listoffigures
\renewcommand{\lstlistlistingname}{Listings}
\lstlistoflistings
\addcontentsline{toc}{chapter}{Listings}
\newpage
\addcontentsline{toc}{chapter}{\nomname}
\Abkuerzung


%%% Chapters
\include{Chapters/introduction}
\include{Chapters/overview}
\include{Chapters/migration}
\include{Chapters/analysis}
%\include{Chapters/application}
\include{Chapters/conclusion}

%%%%%%%%%%%%%%%%%%BIBLIOGRAPHY%%%%%%%%%%%%%%%%%%%%%
%\nocite{*}
\bibliography{bibliography}{}
%\bibliographystyle{apalike}
\bibliographystyle{abbrv}


%\appendix
\begin{appendices}
\include{appendices/xmark-laptop-table}
\include{appendices/xmark-tablet-table}
\include{appendices/xmark-factors-table}
\include{appendices/xmark-tablet-steps-optimized}
\include{appendices/improvements-in-percent}
\include{listings/statement1}
\end{appendices}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%makeindex thesis.nlo -s nomencl.ist -o thesis.nls
\abbrev{dex}{Dalvik Executable}
\abbrev{XML}{Extensible Markup Language}
\abbrev{VM}{Virtual Machine}
\abbrev{JIT}{Just In Time}
\abbrev{DVM}{Dalvik Virtual Machine}
\abbrev{JVM}{Java Virtual Machine}
\abbrev{APK}{Android Application Package}
\abbrev{IPC}{Inter-process Communication}
\abbrev{SDK}{Software Development Kit}
\abbrev{GUI}{Graphical User Interface}
\abbrev{XML}{Extensible Markup Language}
\abbrev{ADT}{Android Development Tools}
\abbrev{HTML}{Hypertext Markup Language}
\abbrev{ART}{Android Runtime}
\abbrev{DOM}{Document Object Model}
\abbrev{SAX}{Simple API for XML}
\abbrev{API}{Application Programming Interface}
\abbrev{JNI}{Java Native Interface}
\abbrev{JDK}{Java Development Kit}




\end{document}

